results

### 2020-11-18 analysis/results.Rmd
funder charaterusutucs added

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
```

@@

sponsor_country <- hybrid_invoice %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  group_by(country_factor) %>%
  summarise(n = n_distinct(doi)) %>%
  mutate(prop = n / sum(n))
```

Figure presents a breakdown by country, highlighting the dominating role of a few countries. Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "Multinational") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%). In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016. The prominent role of Norway can be also explained by a national deal, which started in 2019. US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).

```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)'}
sponsor_country %>%
  mutate(country_factor = fct_reorder(country_factor, n)) %>%
  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
) %>%
  ggplot(aes(country_factor, n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::number_format(big.mark = " ")) +
  coord_flip() +
  theme_minimal_vgrid() +
  labs(y = "Number of invoiced OA articles", x = NULL)
```

>>>>

```{r}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
```

@@

sponsor_country <- hybrid_invoice %>%
  # manual changes
  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2))

sponsor_country %>% 
  ungroup() %>%
  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
  janitor::adorn_totals(c("col", "row")) %>%
  rename(`Country/Subject Area` = country_factor) %>%
  knitr::kable()
```

Figure presents a breakdown by country, highlighting the dominating role of a few countries. Not surprisingly, mostly British funders paid for hybrid open access in Elsevier hybrid journals (`r round(sponsor_country %>% filter(country_factor == "United Kingdom") %>% pull(prop) * 100, 0)`%), followed by funding bodies from the European Commission (`r round(sponsor_country %>% filter(country_factor == "EU") %>% pull(prop) * 100, 0)`%) and the Netherlands (`r round(sponsor_country %>% filter(country_factor == "The Netherlands") %>% pull(prop) * 100, 0)`%). In the latter case, invoices for individual open access articles were not send to research funders, but to the Dutch VSNU, a consortium  of Universities in the Netherlands that has negotiated agreements with Elsevier since 2016. The prominent role of Norway can be also explained by a national deal, which started in 2019. US-based open access sponsors originated mainly from the Bill & Melinda Gates Foundation (n = `r hybrid_invoice %>% filter(oa_sponsor_name == "Bill & Melinda Gates Foundation") %>% distinct(doi) %>% nrow()`).

```{r fig.cap  = 'Invoiced sponsors of open access articles in Elsevier hybrid journals by country (2015-19)', eval = FALSE}
sponsor_country %>%
  mutate(country_factor = fct_reorder(country_factor, n)) %>%
  mutate(country_factor = forcats::fct_relevel(country_factor, "Other", after = 0L)
) %>%
  ggplot(aes(country_factor, n)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::number_format(big.mark = " ")) +
  coord_flip() +
  theme_minimal_vgrid() +
  labs(y = "Number of invoiced OA articles", x = NULL)
```
